A weakly structured stem for human origins in Africa

### 2023-01-14 paper.tex
small edits

We also find that Stem 2 continued to contribute to
western Africans during the Last Glacial Period, indicative that this gene flow
likely occurred in western/central Africa (Table~\ref{tab:migration-rates}).
Such an interpretation is reinforced by the differential migration rates
between regions, i.e. the gene flow from Stem 2 to western Africans is estimated to be 5x that
of the rate to eastern Africans during this period. We performed a variety of
validation tests to explore sensitivity of our assumptions, including relaxing
fixed parameters (SI Section~\ref{sec:validation_strategy}). Most validation
tests resulted in parameters similar to the models discussed above. However,
one exception was the inferred Out of Africa and eastern/western African
divergences which were 10-15ka younger than our fixed parameters. The complexity of
these events warrants the inclusion of additional Eurasian populations and
greater detail in modeling these population expansions; nonetheless, this model
mis-specification does not change other features of our inferred model.

>>>>

We also find that Stem 2 continued to contribute to western Africans during the
Last Glacial Period, indicative that this gene flow likely occurred in
western/central Africa (Table~\ref{tab:migration-rates}).  Such an
interpretation is reinforced by the differential migration rates between
regions, i.e. the gene flow from Stem 2 to western Africans is estimated to be
five times that of the rate to eastern Africans during this period. We
performed a variety of validation tests to explore sensitivity of our
assumptions, including relaxing fixed parameters (SI
Section~\ref{sec:validation_strategy}). Most validation tests resulted in
parameters similar to the models discussed above. However, one exception was
the inferred Out of Africa and eastern/western African divergences which were
10-15ka younger than our fixed parameters, which seem at odds with the
archaeological record \citep{Hublin2020-nk,Hajdinjak2021-jo} and some
genetic estimates \citep{Schlebusch2017-nl,Kamm2020-vn}. The likely
complexity of the demographic events that we model as population splits
warrants the inclusion of additional Eurasian populations and greater modeling
detail, particularly for the out-of-Africa expansion. However, other features
of our inferred models are weakly affected by this analysis choice.
